Transforming Trash as an Artistic Act

### 2016-01-17 drafts/chapter1.tex
Relocation, turnitin fixes
Former-commit-id: e8de078896595446f38348c7c970a0b9f22fae30

Common perception to the trash is to get rid of it as soon as possible. People want to depose trash from their lives most of time not thinking beyond. However to establish solid understanding of trash, we have to give attention to the trash and ask why it is trash? How does it become trash?

@@

In this thesis work to understand the different aspects of trash is a key element. \todo[inline]{These aspects are \ldots} As scholars agreed on \todo{ref.} it is part of our life and daily practice and it is very common concepts from developed cities to rural areas, from modern societies to ancient societies. Sometimes it is accepted as a problem (waste crises, ecological problems.) that is need to carefully and seriously managed. On the other hand it is accepted as a source of diversity. (To establish a solid understanding for trash, it is important to see its dilemmas (it is really a dilemma?) or aspects). Therefore we should ask these questions: What is trash? How does it became trash? Is it a end product or a source material? How much is it valuable? How much is it dangerous?

@@

%\comment{Dört bir yanımız metalarla çevrilmş durumda, istediğimiz şeylere çok kolay bir şekilde ulaşabiliyoruz. Yemek siparişi veriyoruz ayağımıza geliyor. İki blok ötedeki kahveciden istediğimiz zaman kahvemizi alıyoruz. Ama tüm bunların bir bedeli var. Bir yandan da sürekli olarak çöp üretiyoruz. Her ne kadar son derece sistemli bir şekilde çöpler toplanıyor ve belki de ne kadar tüketildiğini hiçbir zaman farkına var mıyoruz?}

% Trash is everywhere and, produced every time
Modern (developed) societies are continuously generating trash and, pile them on landfills. During daily activities (drinking, eating, waiting in the bank?) trash is generated and people get rid of them as quickly as possible. (Various objects become trash after their primary functions consumed. People do not care the package of the objects that they buy. They buy the coffee not the cup of it. After coffee finished the life of cup also finishes. Very small life. But is that really so? Is that really life of an object ends after tossed on the waste bin? \todo{Rubbish Theory, Zizek} (Objects life time in the nature is more than human being. People use them for just 5minutes, however they will exist years on the nature.) The vast amount of industrial discarded items spread through the landfills to oceans. They are the result of highly complex industrial production methods. \comment{They are not easily disposable items in the nature.} They live in the nature thousands of years. They are durable(in terms of resistant to natural affects) products. Most of them packages that are used to carry or protect other materials. After real material used these packages became valueless (or useless). (types of trash can be mentioned here, but currently in the artwork I'm using paper packages, therefore, it is more important.) How manage the all this increasing trash that damaging nature?  This is the common approach to trash and the main problem. (actually the sustainability problem.) It is not the only problem, It can be thought that it is a losing the ability to transform new things, alternative behaviors etc. (Instead of creating new opportunities or alternatives, it is a consuming all of them (which?) and producing huge pile of trash.)

@@

People tend to think that trash is valueless because it is trashed no longer needed or not wanted anymore. However is it really values? What does make it valueless?

People do not think what happens after they throw it away. It exits from your life but not the world. What happens when you throw it away? It is stacked from another place. Value of objects are not static. Trash is not static \todo{ref. first paragraph.} also it has a long journey. A project conducted by MIT researches journey of trash by placing trackers onto the trash \cite{chen2009mit}. The results are surprising. Trashes spread away across the country and this journey takes month. \todo[inline]{give exact reference} We know that it is not limited with the border of America because it is the one of the countries export their garbage to the other countries --- fourth world countries\todo[inline]{footnote what are these countries?}. In other worlds Americans trash is not only their trash. As commodities spread to the every corner of the world, trash also. What happens when it is traveled to the other countries, how they approaches these items. \comment{They find new uses and meanings on them.} \todo[inline]{consider Papua genen people}

\begin{quote}
\paraphrase{“Nobody wonders where, each day, they carry their load of refuse. Outside the city, surely; but each year the city expands, and the street cleaners have to fall farther back. The bulk of the outflow increases and the piles rise higher, become stratified, extend over a wider perimeter” – Italo Calvino, Invisible Cities} The project is an initial investigation into understanding the 'removal-chain' in urban areas and it represents a type of change that is taking place in cities: a bottom-up approach to managing resources and promoting behavioral change through pervasive technologies. TrashTrack builds on previous work of the SENSEable City Lab in its exploration of how the increasing deployment of sensors and mobile technologies radically transforms how we understand and describe cities. “People just take their trash and put it on the curb and they forget about it and don’t think about all the time and energy and money put into disposing of it.”
\end{quote}

% Cycle of trash
Trash moves, objects moves from place to place. From homes to garbage trucks. From streets to land fills. From garbage baskets to sculptures. \todo[inline]{REF. MIT Garbage project} Lots of different people touches to trash. With the trash what moves?

% ARTWORK http://timgaudreau.com/2012/trash/trash.html
[ARTWORK] \comment{Self-Portrait as Revealed by Trash} Even if people pay little attention, everybody do not. An artist takes photos of his every trash, like people take photos of their faces. He focuses on the thing that paid little attention, that is trash. At the end he exhibits all the photos by covering all the walls of room. It can seen as record of his activities, record of commodities. Most of them we are not aware of them. Draw the attention of is being thrown away and every time and everywhere. They are too much and can only fit covering all the walls. \todo[inline]{Tim Gaudreau “Self-Portrait as Revealed by Trash: 365 days of photographing everything I threw out – Variation I, 2006”} \paraphrase{After photographing my trash over the course of the year, I ended up with 5,000 images representing my waste stream. These collections of grids represent my first experiments with using those images. This initial work ultimately led to the full scale collages that have become my Self-portrait as Revealed by Trash: 365 days of photographing everything I threw out.}

% ARTWORK http://www.keaggy.com/chairs/sad/12/
\todo[inline]{Sad chair. Photos of lonely chairs that are no longer wanted. Artist captures the continuously left alone chairs.}

Why people are interested in rubbish. Some of the artist also interested in others trash. \comment{O çöplerde ne bulmayı umuyorlar ki? Aslında çöp bizim davranışlarımız, yaşayız tarzımız hakkında bize yön veriyor. ve bu her zaman fiziksel dönüştürme ile değil de aynı zamanda anlamsal ve mekansal değiştirme sorulamalarla mümkün oluyor. Bazen de olduğu gibi kullanıyorlar. bu açıdan önemli. Çöpü çöp şekliden sunan sanatçılar. Ama bu nasıl mümkün ki? fotoğrafını çekenler aslında gerçekten ona müdahele etmemiş mi oluyorlar. Bence değil? bir şekilde müdahele var gene orada. o fotoğraflarda artık dikkatin objesi haline geliyor. çöpün dönüşmesi aslında bir diğer yandan çöpe bakışın değişmesi olarak görülebilir.}

% TODO Life time of objects.
% Nesnelerin yaşam süreleri... Can you imagine how much plastic cup is being consumed every hour?

@@

%****************************************
% EVERYTHING CAN BE TRASH. Act of trashing not act of transforming. Trash is like kara delik gibi ışığın odan kaçamaması gibi hiç bir şeyin ondan kaçılmayacağını hissediyor aslında. 
% Michael Landy's Art Bin uses a art gallery to create his dust bin. Describing the work, simply called Art Bin, as 'about failure', Landy is inviting members of the public to bring their own artistic failures along to the gallery from 29 January, where their worthlessness will be assessed. Damien Hirst, Gillian Wearing, Tracey Emin and Mark Titchner have already contributed, offering sculptures, paintings and prints. "There's no hierarchy once they are in the bin" All of them are accepted as same. Ultimate equality. Tosing them to the dust bin makes them rubbish even if they are combined from failed attempts. Nothing's too good for the art bin. Everything can fit into the art bin. Michael Landy transforms the SLG into Art Bin, a container for the disposal of works of art. As people discard their art works the enormous 600m³ bin becomes, in Michael Landy’s words, “a monument to creative failure”. It is a very big glass bin and isolates people. There is a provocative approach to the art and art objects by saying that "Nothing's too good for the art bin". There is no limitation of throwing away even if art.

% Herkes çöpü dönüştürmüyor, bazıları ise çöpe dönüştürüyor. Ama ben bununla ilgilenmiyorum aslında.
%........................................

\todo[inline]{ArtBin, squeezing. [Trashing Things, Artwork]}

%http://www.smithsonianmag.com/arts-culture/photographer-creates-fine-art-out-trash-we-throw-environment-180951615/?no-ist
\todo[inline]{showing trash as they are... [Artwork] by Barry Rosenthal}

%****************************************
% Different types of trash
\todo[inline]{Different types of trash.} \comment{There are many works from trash around the world. Build up trash. Collected from graveyards, landfills, unused materials etc.}

The complexity of produced trashes of societies is increasing. For example developed countries that have nuclear plant generates radioactive wastes which highly hazardous for the environment is never exist previous societies. Think batteries and so on. Every society generates different types of wastes. Differs from country to country, society to society, ages to ages. It can be thought that when the complexity of trashed increased required effort to repair, reuse and recycle is also increase. Therefore for the ones that have no complex tools it is becoming harder to reuse objects. In other words objects become more complex their re-usage becomes less likely.  Different production process generates different types of trash. According to production process, decomposition process\ldots The approach to the different type of trash will be different. In other word if trash is a result of classification of objects, it can be easily extracted that there is classification inside of it. There are some trashes that are more close to the people. More easy to convert them. more easy to regain to the society. 
%........................................

%****************************************
% Different disciplines, Perspectives related with trash from different disciplines
\todo[inline]{List of disciplines, and their approach to the trash.} \comment{Trash draws attention different peoples and disciplines. In other words there are different studies related with trash. Different approaches to the trash and son on.}

@@

As many people are inside the business of transforming trash through craft and so on. People on the streets looking for aluminum cans, glass bottles, plastic. They collect and sell them to survive. There are companies that are collect waste materials to regain them again the industry. In thesis they are not in the scope. People that give attention to the trash and behave artistic act is in the scope of this thesis.

\comment{How does it become trash?} Here the purpose is to understand the dynamics that turn objects to trash. By understanding them is provide a road map (or ideas) how to turn trash to something valuable? (The purpose of this thesis is to find (or explore) a way(methodology, approach) to add value to object using artistic methods? Therefore first question is why they are less-valued, and ignored. How to make them valuable? How to make them part of our life again?)

%%%
%%%
%%% 
\section{Structure of the Study}
This study is structured by many examples of artworks in different part of the thesis. When a topic or argument is presented, immediately example artwork is given at this moment and how artist replies this issue is illustrated. Thereby artworks are examined in the theoretical and cultural context. Artworks are used as a supportive elements for the stated arguments.  However, not every time artworks are discussed in detail. Only basic information is given and how it is related with this topic explained. \comment{By the way} it is better to keep in mind that all works can be analyzed in different contexts in more detail fashion.

@@

%%%
%%%
%%%
\section{A Note on Terminology}
Within the scope of thesis only words that are English is examined but in different languages and cultures there might be different naming for various stages of object and trash. Analysis of them may be subject of another thesis. (Garbage, trash, rubbish, debris, detritus, waste, scrap, junk, refuse, discard, disposal, litter: during my research scholars and authors have different names for the stuff.) I realized that many scholars and resources have used different words (or terminology) to define trash. Although there are slight differences between these words, sometimes they are used to signify same concept (or used with same purposes). Understanding diverse vocabulary of this topic before dive into the theory makes more clearer the borders of the work. There is no consensus on terminology. Some scholars uses junk art, others recycling art. The origin of vocabulary of this topic reveals a whole side of human activity: our history revealed by what we have thrown away through the ages. What were people throwing out when these words were coined? From the rubbish: the archeology of garbage gives clearer definition of some these words:

@@

@@

%****************************************
% Reuse and Recycle
One of the most encountered words in this topic (transforming trash) is reuse and recycle. Both of them are used widely by scholars \todo{ref}

@@

@@

@@

In this thesis scope mainly single use, disposable objects are considered generally. Nearly half of the material used in the project is disposable packets. Therefore as well as other words used but mainly trash is used to refer the material.

Some of the scholar uses recycling\cite{cerny1996recycled,herman1998trashformations} for the recreation and transformation of it. Some of the scholars and artist uses upcycling for their process.

>>>>

\paraphrase{The phenomenon of waste comes clearly into focus not merely as a by-product of manufacturing processes, but rather as an integral element in cycles of production and consumption. (Trash Culture)}

\comment{Common approach on trash is dominated by the environmentalist.}

\comment{Last decades much attention has been devoted to the waste.}

\comment{At least since the early twentieth century, the concern with discarded things and materials has been a recurring theme in art.} \todo{Bu nereden?}

\paraphrase{Over the course of the twentieth century, the twin developments of mass production and mass media in the capitalist economies of the Global North completed a total transformation of everyday life, reorienting almost every activity toward consumption. Things once locally produced and often handmade were now mass produced and commodified, turning local, artisanal producers into deskilled laborers serving the assembly line. This world of mass production radically altered the meaning of objects in an unprecedented and profound reification that reached into every sphere of life. This was not just the fate of objects, but also of the practices with which they are always enmeshed. Mass production and commodification liquidated what rema ined of folkways tied to local production, demanding people construct the meaning of their lives through purchases rather than production. This transition to a lifeworld of consumption affected not only concrete objects, but even more forcefully altered symbolic and aesthetic practices. Technologies of mechanical reproduction redefine d story, image, and music, altering the traditions of both fine and fol k art with mass produced and distributed forms in newspapers, advertising, radio, film, and television that shattered the aura of fine art and liquidated folk art almost completely. Rather than making their stories, images, and music, ever more urbanized workers and manage rs consumed new massproduced art forms that, epitomized b y the Hollywood studio system, developed into the walltowall media scape of twentyfour hour broadband that now blankets the city, the suburb, and the country alike. (FROM Collage Culture)}

\paraphrase{Given these developments, this book seeks to answer the following questions: How did art and literature respond to this age of consumption? What do the productions and practices of artists and writers reveal about the meaning of mass production, consumption, reification, mechanical reproduction, and meaning? (FROM Collage Culture)}

[COMMON PERCEPTION] (In the dictionary trash means anything useless, worthless or discarded. In parallel to this definition) Common perception to the trash is to get rid of it as soon as possible. People want to depose trash from their lives most of time not thinking beyond. However to establish solid understanding of trash, we have to give attention to the trash and ask why it is trash? How does it become trash? People tend to think that trash is valueless because it is trashed no longer needed or not wanted anymore. However is it really values? What does make it valueless?

@@

It is part of our consumption activities (/practices) and it is very common concepts from developed cities to rural areas, from modern societies to ancient societies.

[DILEMMA] In this thesis work to understand the different aspects of trash is a key element. \todo{These aspects are \ldots} Sometimes it is accepted as a problem (waste crises, ecological problems.) that is need to carefully and seriously managed. On the other hand it is accepted as a source of diversity. (To establish a solid understanding for trash, it is important to see its dilemmas (it is really a dilemma?) or aspects). Therefore we should ask these questions: What is trash? How does it became trash? Is it a end product or a source material? How much is it valuable? How much is it dangerous?

@@

% Trash is everywhere and, produced every time
[LIFE OF OBJECTS] Modern (developed) societies are continuously generating trash and, pile them on landfills. During daily activities (drinking, eating, waiting in the bank?) trash is generated and people get rid of them as quickly as possible. (Various objects become trash after their primary functions consumed. People do not care the package of the objects that they buy. They buy the coffee not the cup of it. After coffee finished the life of cup also finishes. Very small life. But is that really so? Is that really life of an object ends after tossed on the waste bin? \todo{Rubbish Theory, Zizek} (Objects life time in the nature is more than human being. People use them for just 5minutes, however they will exist years on the nature.) The vast amount of industrial discarded items spread through the landfills to oceans. They are the result of highly complex industrial production methods. \comment{They are not easily disposable items in the nature.} They live in the nature thousands of years. They are durable(in terms of resistant to natural affects) products. Most of them packages that are used to carry or protect other materials. After real material used these packages became valueless (or useless). (types of trash can be mentioned here, but currently in the artwork I'm using paper packages, therefore, it is more important.) How manage the all this increasing trash that damaging nature?  This is the common approach to trash and the main problem. (actually the sustainability problem.) It is not the only problem, It can be thought that it is a losing the ability to transform new things, alternative behaviors etc. (Instead of creating new opportunities or alternatives, it is a consuming all of them (which?) and producing huge pile of trash.)

@@

[MIT, COMMON APPROACH] People do not think what happens after they throw it away. It exits from people's life but not the world. What happens when you throw it away? It is stacked from another place. \comment{Value of objects are not static.} Trash is not static \todo{ref. first paragraph.} also it has a long journey. A project conducted by MIT researches journey of trash by placing trackers onto the trash \cite{chen2009mit}. The results are surprising. Trashes spread away across the country and this journey takes month. Further it is not limited with the border of America because it is the one of the countries export their garbage to the other countries --- fourth world countries\todo[inline]{footnote what are these countries?}. In other words American's trash is not only their trash. As commodities spread to the every corner of the world, trash also. What happens when it is traveled to the other countries, how they approaches these items. \comment{They find new uses and meanings on them.} \todo[inline]{consider African native}

\begin{quote}
\paraphrase{Nobody wonders where, each day, they carry their load of refuse. Outside the city, surely; but each year the city expands, and the street cleaners have to fall farther back. The bulk of the outflow increases and the piles rise higher, become stratified, extend over a wider perimeter (by Italo Calvino, Invisible Citie). People just take their trash and put it on the curb and they forget about it and don’t think about all the time and energy and money put into disposing of it.}
\end{quote}

\comment{MIT project can be seen as a different approach to understand the trash and its \quotes{removal-chain} from people's life.}

\comment{Biodegradable matter is material capable of being decomposed by bacteria or other biological means. Biodegradable  matter usually consists of organic materials such as plant, animal, and other substance matter originating from living  organisms. It has the ability to be broken down into smaller, harmless products by way of the action of living organisms.  The term is often used in relation to ecology, garbology, and waste management. Biodegradable products are often associated with perishables (products, food, and waste materials subject to death or decay). The term biodegradability of a product  refers to its disposition to disintegrate as the result of natural processes.}

% Cycle of trash
[MOVES] Trash moves, objects moves from place to place. From homes to garbage trucks. From streets to land fills. From garbage baskets to sculptures. Lots of different people touches to trash. With the trash what moves? Trash moves around. Travels like humans. Travels like commodities. When we are telling about the travel of commodities, then it is easy to say that trashes also moves around.

[IMPORTANT, QUESTINING] Why people are interested in rubbish. Some of the artist also interested in others trash. \comment{O çöplerde ne bulmayı umuyorlar ki? Aslında çöp bizim davranışlarımız, yaşayız tarzımız hakkında bize yön veriyor. ve bu her zaman fiziksel dönüştürme ile değil de aynı zamanda anlamsal ve mekansal değiştirme sorulamalarla mümkün oluyor. Bazen de olduğu gibi kullanıyorlar. bu açıdan önemli. Çöpü çöp şekliden sunan sanatçılar. Ama bu nasıl mümkün ki? fotoğrafını çekenler aslında gerçekten ona müdahele etmemiş mi oluyorlar. Bence değil? bir şekilde müdahele var gene orada. o fotoğraflarda artık dikkatin objesi haline geliyor. çöpün dönüşmesi aslında bir diğer yandan çöpe bakışın değişmesi olarak görülebilir.}

@@

%****************************************
% Different types of trash
[Different types of trash] The complexity of produced trashes of societies is increasing. For example developed countries that have nuclear plant generates radioactive wastes which highly hazardous for the environment is never exist previous societies. Think batteries and so on. Every society generates different types of wastes. Differs from country to country, society to society, ages to ages. It can be thought that when the complexity of trashed increased required effort to repair, reuse and recycle is also increase. Therefore for the ones that have no complex tools it is becoming harder to reuse objects. In other words objects become more complex their re-usage becomes less likely.  Different production process generates different types of trash. According to production process, decomposition process\ldots The approach to the different type of trash will be different. In other word if trash is a result of classification of objects, it can be easily extracted that there is classification inside of it. There are some trashes that are more close to the people. More easy to convert them. more easy to regain to the society.

\comment{Biodegradable waste, Human waste, Organic waste, Food waste, Biomedical waste, Clinical waste, Medical waste, Household waste, Radioactive waste, Marine debris, Electronic waste (e-waste), Construction and demolition waste\ldots}

\comment{Liquid type, Solid type, Hazardous type, Organic type, Recyclable type}

\comment{Radioactive waste is waste that contains radioactive material. Radioactive waste is usually a by-product of nuclear power generation and other applications of nuclear fission or nuclear technology, such as research and medicine. Radioactive waste is hazardous to most forms of life and the environment, and is regulated by government agencies in order to protect human health and the environment.}
%........................................

%****************************************
% Different disciplines, Perspectives related with trash from different disciplines
\comment{Trash draws attention different peoples and disciplines. In other words there are different studies related with trash. Different approaches to the trash and son on.}

trash have influenced, and are also influenced by, cultural products such as films, visual art, museum exhibits and literature.

@@

[Different approaches] As many people are inside the business of transforming trash through craft and so on. People on the streets looking for aluminum cans, glass bottles, plastic. They collect and sell them to survive. There are companies that are collect waste materials to regain them again the industry. In thesis they are not in the scope. People that give attention to the trash and behave artistic act is in the scope of this thesis.

In the scope of trash not biodegradable trash is considered. Produced, industrial commodities are considered.

%%%
%%%
%%% 
\section{Structure of the Study}
\comment{Theory and artworks are not strictly separated from each other. Artworks are explained with the help of theories.}

\comment{Research and the found samples of works helped a lot to shape my work. They provide me deeper understanding of the subject and what I am doing and they lead me during the development of the work.}

\comment{the thesis is separated into three part.}

This study is structured by many examples of artworks in different part of the thesis. When a topic or argument is presented, immediately example artwork is given at this moment and how artist replies this issue is illustrated. Thereby artworks are examined in the theoretical and cultural context. Artworks are used as a supportive elements for the stated arguments.  However, not every time artworks are discussed in detail. Only basic information is given and how it is related with this topic explained. \comment{By the way} it is better to keep in mind that all works can be analyzed in different contexts in more detail fashion.

@@

%%%
%%%
%%%
\section{A Note on Terminology}
\comment{using words interchangeably.}

Within the scope of thesis only words that are English is examined but in different languages and cultures there might be different naming for various stages of object and trash. Analysis of them may be subject of another thesis. (Garbage, trash, rubbish, debris, detritus, waste, scrap, junk, refuse, discard, disposal, litter: during my research scholars and authors have different names for the stuff.) I realized that many scholars and resources have used different words (or terminology) to define trash. Although there are slight differences between these words, sometimes they are used to signify same concept (or used with same purposes). Understanding diverse vocabulary of this topic before dive into the theory makes more clearer the borders of the work. There is no consensus on terminology. Some scholars uses junk art, others recycling art. The origin of vocabulary of this topic reveals a whole side of human activity: our history revealed by what we have thrown away through the ages. What were people throwing out when these words were coined? From the rubbish: the archeology of garbage gives clearer definition of some these words:

@@

\paraphrase{A note on terminology: The chapters in this book feature a broad range of   terms, reflecting the multifaceted and conceptually complex nature of   the issue at hand. Many of   these terms – such as ‘trash’, ‘garbage’ and ‘rubbish’ – are frequently used interchangeably. This is not least because the variety of   terms reflects the variations in English and US vocabulary and thus all usages are retained as individual authors intended in order to reflect the international nature of   this volume. Other related terms, such as ‘ruins’, ‘obsolete’, ‘waste’, ‘discards’ are used in a variety of contexts in each chapter. Again, these usages have been retained to reflect the focus of individual studies. (Trash Culture)}

@@

%****************************************
% Reuse and Recycle
One of the most encountered words in this topic (transforming trash) is reuse and recycle. Both of them are used widely by scholars. \todo{ref}

@@

\comment{There is a difference between garbage and waste, although garbage is a part and an aspect of waste.}

@@

Some of the scholar uses recycling \cite{cerny1996recycled,herman1998trashformations} for the recreation and transformation of it. Some of the scholars and artist uses upcycling for their process. \todo{move to the more appropriate place}

@@
